Squared grid puzzle

### 2021-03-13 Paper_Squared_Grid.tex
Completed title, abstract and first section of the paper
Added new explanatory pictures.

	% Figures and styles
	\usepackage[Export]{adjustbox} % Used to constrain images to a maximum size
	\usepackage{float}
	\floatplacement{figure}{H} % forces figures to be placed at the correct location
	\usepackage{graphicx} % Basic figure setup
	\usepackage{fancyvrb} % verbatim replacement that allows latex
	\usepackage{grffile} % extends the file name processing of package graphics

@@

	%% TODO this should be title and subtitles
	\section{Squared grid puzzle}
	\subsubsection{An informal analysis of the classical nine-dots puzzle	``under steroids''}

	\begin{figure}[H]
		\centering
		\includegraphics[width=0.5\textwidth]{images/egg_puzzle.jpg}
		\caption{The original illustration of the nine-dots puzzle in the Sam Loyd's "Cyclopedia of puzzle" (1914)}
	\end{figure}

	\hypertarget{introduction} {
		\section{Introduction}
		\label{introduction}
	}
	The \textbf{nine dots puzzle} has been first cited by Sam Loyd's in its
	1914 \emph{Cyclopedia of Puzzles} (see image above). This intriguing
	conundrum was later revived by psychologists, salesmen and ``business
	coaches'' becoming a mainstream \textbf{``thinking-outside-the-box''}
	brain-teaser, i.e.~a benchmark of creative divergent thinking.

	In its basic formulation, the goal is to join nine dots arranged in a
	squared grid with the fewest number of straight lines, and without
	lifting the pen.
	\begin{figure}[H]
	\includegraphics[width=0.2\textwidth]{images/9-dots-grid.png}
	\includegraphics[width=0.2\textwidth]{images/9-dots-solution.png}
	\end{figure}
	The basic version of the puzzle is very famous so you can easily learn
	more about it by searching online, e.g.~you can start from the
	\href{https://en.wikipedia.org/wiki/Thinking_outside_the_box}{Wikipedia
	page}.

	In this paper, I'll discuss a generalized \(2D\) version of the original
	problem, which is extended to a squared grid with an arbitrary number of
	points: the starting \emph{9-dots puzzle} is now the generic
	\emph{9-16-25-36\ldots{} dots puzzle}. First, I'll show that the general
	solution can be drawn iteratively, as one may have already guessed. You
	will also be stunned to learn that, starting from a particular grid size
	upwards, salesmen can no longer refer to the grid problem as a paradigm
	of ``out-of-the-box thinking''. Last but not least, I propose an
	intriguing arithmetical interpretation of the solution for the general
	bidimensional problem.

	\textbf{Disclaimer}: the squared grid problem has already been solved
	and thoroughly treated. Marco Ripà and Pablo Ramirez first extended the
	nine dots problem to a generic \(2D\) squared grid and then also to
	multiple dimensions! You can find their papers in the \hyperlink{references}{References} section.\\
	This document simply aims to be an enjoyable and informal digression on
	the squared grid problem as posed by the aforementioned authors. As
	regards the arithmetical interpretation of the solution to the
	bidimensional puzzle that I'll provide, so far I didn't find any similar
	counterpart by searching online but I may be wrong: in case it's an
	original remark, you're welcome!

	\hypertarget{the-extended-problem-and-its-solution} {
		\section{The extended problem and its solution}
		\label{the-extended-problem-and-its-solution}
	}

@@

	\paragraph{Acknowledgements} \mbox{} \\ % mbox needed for having carriage return after "paragraph"
	Marco Ripà posed the squared grid brain-teaser during one of its
	stimulating YouTube live sessions. I am obliged to him for the great fun
	I had in tackling this problem, which in turn inspired all the analysis
	and ideas that followed.

	\paragraph{Technical note} \mbox{} \\
	Python was used to code the generic solver algorithm and produce the
	explanatory images of the paper. You can find the implementation code in
	the attached Jupyter notebook, in which you can also display and save a
	nice animation drawing the solution to the puzzle.

	\paragraph{Future developments} \mbox{} \\
	I'll work on solving the problem in multiple dimensions, you bet! I
	would also like to extend the solver's code to handle this more general
	case; hence, as a plus I could produce animations drawing the solution
	for three-dimensional grids.

@@

	\begin{itemize}
		\item
			The basic nine-dots puzzle has been first cited here: \emph{Cyclopedia of Puzzles, Samuel Loyd (1914)}
		\item
			A link to a really valid \href{http://utenti.quipo.it/base5/geopiana/loyd9punti4linee.htm}{online reference (in Italian)}
		\item
			\emph{Marco Ripà and Pablo Ramirez, Il Nine Dots Puzzle esteso a N×N×\ldots×N punti}.
		\item
			\emph{Marco Ripà, Nine Dots Puzzle extended to N1×N2×\ldots×Nk points 	under house arrest}.
	\end{itemize}

>>>>

% Figures and styles
%\usepackage[Export]{adjustbox} % Used to constrain images to a maximum size
\usepackage{float}
%\floatplacement{figure}{H} % forces figures to be placed at the correct location
\usepackage{graphicx} % Basic figure setup
%\usepackage{fancyvrb} % verbatim replacement that allows latex
\usepackage{grffile} % extends the file name processing of package graphics

@@

\newcommand{\subtitle}[1]{%
  \posttitle{%
    \par\end{center}
    \begin{center}\large#1\end{center}
    \vskip0.5em}%
}

\title{Squared grid puzzle}
\subtitle{An informal analysis of the classical nine-dots puzzle ``under steroids''}
\date{}
\author{\\ Andrea Rebolini}

\maketitle

\begin{figure}[H]
\centering
\includegraphics[width=0.4\textwidth]{images/think-out-the-box-icon.jpg}
\end{figure}

\begin{abstract}
The nine-dots puzzle is a mainstream brain-teaser. It's often sponsored by business coaches and salesmen as a benchmark of creative thinking, as its solution requires to ``think outside-the-box".
In this paper, I review the nine-dots puzzle and then analyze a generalized version of the original problem, which is extended to a squared grid consisting of an arbitrary number of points: i.e, the starting \emph{9-dots puzzle} becomes the generic \emph{9-16-25-36.. dots puzzle}. First, I show how the general solution can be achieved. You'll then learn that starting from a particular grid size upwards, salesmen can no longer enthrall you with out-of-the-box tricks. Finally, I propose an intriguing arithmetical interpretation of the solution to the general bidimensional problem.
\\
\\
\textbf{Disclaimer}: the squared grid problem has already been solved and thoroughly treated. Marco Ripà and Pablo Ramirez first extended the nine-dots puzzle to an arbitrarily-sized squared grid and then also to multiple dimensions! You can find their papers in the \hyperlink{references}{References} section.\\
This document simply aims to be an enjoyable and informal digression on the squared grid problem as posed by the aforementioned authors. As regards the arithmetical interpretation of the solution to the puzzle that I'll provide, so far I didn't find any counterpart by searching online but I may be wrong: in case it's an original remark, you're welcome!
\end{abstract}

\newpage

\hypertarget{nine-dots-puzzle} {
	\section{The nine dots puzzle}
	\label{nine-dots-puzzle}
}
The \textbf{nine-dots puzzle} has been first cited by Sam Loyd's in his \emph{Cyclopedia of Puzzles} (see \autoref{fig:egg_puzzle}).
\begin{figure}[H]
	\centering
	\includegraphics[width=0.54\textwidth]{images/egg_puzzle.jpg}
	\caption{The original illustration of the nine-dots puzzle in the Sam Loyd's "Cyclopedia of puzzle" (1914). The author entitled it as \emph{the Christoforus Columbus egg puzzle}; this is just a reference to the \emph{egg of Columbus}: you don't actually need 9 eggs to try it yourself.}
\label{fig:egg_puzzle}
\end{figure}
In its basic formulation, the goal is to join nine dots arranged in a squared grid with the fewest number of straight lines, and without lifting the pen. Take your time to figure out the solution...
\begin{figure}[H]
\centering
\includegraphics[width=0.3\textwidth]{images/9-dots-grid.png}
\caption{The 9-dots puzzle: join all the points using straight lines, without lifting the pen.}
\label{fig:9-dots-grid}
\end{figure}
If you concluded that at least five lines are required to join all the points, and your solution is something like \autoref{fig:9-dots-wrong} alas! you embraced mediocrity and should not complain about your pen-pusher job. But don't worry: contrast your outcome with the one given by the king in the Sam Loyd's illustration (again \autoref{fig:egg_puzzle}): you're ahead by one line against the king, and thus still highly qualified to rule a nation.
\begin{figure}[H]
\centering
\includegraphics[width=0.3\textwidth]{images/9-dots-grid.png}
\includegraphics[width=0.3\textwidth]{images/9-dots-wrong-solution.png}
\caption{A suspiciously trivial (and thus wrong) solution to the 9-dots puzzle.}
\label{fig:9-dots-wrong}
\end{figure}
How is possible to connect the dots with fewer lines?
What hinders most of the people is that they restrict themselves to draw lines within the grid of points, myopically focusing on an immaginary boundary around the grid (similar to the thin grey borders that I maliciously put in the above figures). In order to get to the ``right" solution (or at least to a more optimal solution: in \hyperlink{appendices}{Appendices}, you'll see why I make such a clarification) you need to \textbf{``think outside the box''}. Using a more creative mindset, a solution can be drawn consisting of just four lines, as shown in \autoref{fig:9-dots-solution}.
%% TODO references to the figure's number are all wrong! But the links are correct.
\begin{figure}[H]
\centering
\includegraphics[width=0.3\textwidth]{images/9-dots-solution.png}
\caption{A creative and effective solution to the 9-dots puzzle. Think outside the box!}
\label{fig:9-dots-solution}
\end{figure}
Nice trick, isn't it?! No surprise, this intriguing conundrum was revived by psychologists, salesmen and business coaches, always seeking a nice \emph{coup-de-théâtre} for their demonstrations.

\hypertarget{the-extended-problem-and-its-solution} {
	\section{The extended problem and its solution}
	\label{the-extended-problem-and-its-solution}
}
%
%First, I'll show that the general solution can be drawn iteratively, as you may have already guessed

@@

\paragraph{Acknowledgements} \mbox{} \\ % mbox needed for having carriage return after "paragraph"
Marco Ripà posed the squared grid brain-teaser during one of its stimulating \href{https://www.youtube.com/watch?v=rh-ONRXcHok}{YouTube live sessions}. I am obliged to him for the great fun I had in tackling this problem, which in turn inspired all the analysis and ideas that followed.

\paragraph{Technical note for computer nerds} \mbox{} \\
Python was used to code the solver algorithm and produce the explanatory images of the paper. You can find the implementation in the attached Jupyter notebook, in which you can also display (and save) a nice animation drawing the solution to the puzzle.

\paragraph{Future developments} \mbox{} \\
I'll address the problem in multiple dimensions, you bet! I would also like to extend the solver's code to handle this more general case; hence, as a plus I could produce animations drawing the solution for three-dimensional grids.

\hypertarget{appendices}{
	\section{Appendices}
	\label{appendices}
}

@@

\begin{itemize}
	\item
		The basic nine-dots puzzle has been first cited here: \emph{Cyclopedia of Puzzles, Samuel Loyd (1914)}.
	\item
	\href{https://en.wikipedia.org/wiki/Thinking_outside_the_box}{Wikipedia page on out-of-the-box thinking} is worth reading: the 9-dots puzzle is thoroughly discussed.
	\item
		A link to a really valid \href{http://utenti.quipo.it/base5/geopiana/loyd9punti4linee.htm}{online reference (in Italian)}
	\item
		\emph{Marco Ripà and Pablo Ramirez, Il Nine Dots Puzzle esteso a N×N×\ldots×N punti}.
	\item
		\emph{Marco Ripà, Nine Dots Puzzle extended to N1×N2×\ldots×Nk points 	under house arrest}.
\end{itemize}
